---
title: "checking_into_geography"
author: "kira_delmore"
date: '`r Sys.Date()`'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

Get things ready
```{r}
setwd("C:/Users/Kira Delmore/Dropbox/Haley and Kira's Comparative Analysis Extravaganza/Analysis")
library(EvoRAG)
library(car)
evodat<-read.csv("../working files/pair_hedges_avg_jul24.csv",stringsAsFactors = FALSE,strip.white = TRUE, na.strings = c("NA",""))

## transform avg variables
evodat["p_distance_100"]<-evodat$p_distance*100
evodat["avg_song10"]<-evodat$avg_song/10
evodat["avg_morph10"]<-evodat$avg_morph/10
evodat["avg_colour10"]<-evodat$avg_colour/10
evodat["avg_all10"]<-evodat$avg_2/10

## set trait for analyses
evodat["trait"]<-evodat$avg_all10 #re-set this for each variable being analyzed
#evodat["trait"]<-evodat$avg_song10
#evodat["trait"]<-evodat$avg_colour10
#evodat["trait"]<-evodat$avg_morph10
```

LIMIT DATASET TO 36 NON-SYM PAIRS AND RUN _NULL MODELS

```{r}
evodat=subset(evodat,evodat$visual_official_jason!="sym")
models = c("BM_null", "OU_null")
```

a) Run model allowing no separate rates
```{r}
e1<-evodat$trait
t1<-evodat$p_distance_100
l1<-evodat$mass_avg #dummy
fit_all<-model.test.sisters(e1,t1,l1,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL) #model for all pairs
#plot(t1,e1,xlim=c(0,10),ylim=c(0,1),main="all")
print(fit_all)
```

b) Run model allowing separate rates
```{r}
## parallels
evodat_parallel<-subset(evodat,evodat$migration_category=="parallel")
e2 <- evodat_parallel$trait
t2<-evodat_parallel$p_distance_100
l2<-evodat_parallel$mass_avg 
fit_par <- model.test.sisters(e2,t2,l2,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t2,e2,xlim=c(0,10),ylim=c(0,1),main="no divide")
print(fit_par) #View model parameters

## perpendiculars
evodat_perp<-subset(evodat,evodat$migration_category=="perpendicular")
e3 <- evodat_perp$trait
t3<-evodat_perp$p_distance_100
l3<-evodat_perp$mass_avg
fit_perp <- model.test.sisters(e3,t3,l3,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t3,e3,xlim=c(0,10),ylim=c(0,1),main="divide")
print(fit_perp) #View model parameters

## calculate AIC for perp/par
logLike_BM<-fit_par[1,1]+fit_perp[1,1]
AIC_BM<-2*1-2*logLike_BM
print(AIC_BM)

logLike_OU<-fit_par[1,2]+fit_perp[1,2]
AIC_OU<-2*2-2*logLike_OU
print(AIC_OU)
```

LIMIT DATASET TO 36 NON-SYM PAIRS AND RUN _LINEAR* MODELS

```{r}
evodat$migration_category<-recode(evodat$migration_category,"'perpendicular'=0;'parallel'=1")
models = c("BM_null", "BM_linear", "OU_null", "OU_linear","OU_linear_beta")
```

a) Run model allowing no separate rates
```{r}
e1<-evodat$trait
t1<-evodat$p_distance_100
l1<-evodat$mass_avg #dummy
fit_all<-model.test.sisters(e1,t1,l1,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL) #model for all pairs
#plot(t1,e1,xlim=c(0,10),ylim=c(0,1),main="all")
print(fit_all)
```

b) Run model allowing separate rates
```{r}
e9 <- evodat$trait
t9<-evodat$p_distance_100
g9<-evodat$migration_category
fit_par_patry <- model.test.sisters(e9,t9,g9,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t9,e9,xlim=c(0,10),ylim=c(0,1),main="no divide")
print(fit_par_patry) #View model parameters
```

GO BACK TO FULL DATASET AND USE OVERLAP PROPORTION TO EXAMINE GEOGRAPHIES INFLUENCE

```{r}
setwd("C:/Users/Kira Delmore/Dropbox/Haley and Kira's Comparative Analysis Extravaganza/Analysis")
evodat<-read.csv("../working files/pair_hedges_avg_jul24.csv",stringsAsFactors = FALSE,strip.white = TRUE, na.strings = c("NA",""))

## transforming avg variables
evodat["p_distance_100"]<-evodat$p_distance*100
evodat["avg_song10"]<-evodat$avg_song/10
evodat["avg_morph10"]<-evodat$avg_morph/10
evodat["avg_colour10"]<-evodat$avg_colour/10
evodat["avg_all10"]<-evodat$avg_2/10

## set trait for all analyses
evodat["trait"]<-evodat$avg_all10 #re-set this for each variable being analyzed
#evodat["trait"]<-evodat$avg_song10
#evodat["trait"]<-evodat$avg_colour10
#evodat["trait"]<-evodat$avg_morph10
```

a) Run model allowing no separate rates (same as first set of models but with full dataset)
```{r}
models = c("BM_null", "OU_null")

e1<-evodat$trait
t1<-evodat$p_distance_100
l1<-evodat$mass_avg #dummy
fit_all<-model.test.sisters(e1,t1,l1,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL) #model for all pairs
#plot(t1,e1,xlim=c(0,10),ylim=c(0,1),main="all")
print(fit_all)
```

b) Run model allowing separate rates (again, same as first set of models but with full dataset)
```{r}
## parallels
evodat_parallel<-subset(evodat,evodat$migration_category=="parallel")
e2 <- evodat_parallel$trait
t2<-evodat_parallel$p_distance_100
l2<-evodat_parallel$mass_avg 
fit_par <- model.test.sisters(e2,t2,l2,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t2,e2,xlim=c(0,10),ylim=c(0,1),main="no divide")
print(fit_par) #View model parameters

## perpendiculars
evodat_perp<-subset(evodat,evodat$migration_category=="perpendicular")
e3 <- evodat_perp$trait
t3<-evodat_perp$p_distance_100
l3<-evodat_perp$mass_avg
fit_perp <- model.test.sisters(e3,t3,l3,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t3,e3,xlim=c(0,10),ylim=c(0,1),main="divide")
print(fit_perp) #View model parameters

## calculate AIC for perp/par
logLike_BM<-fit_par[1,1]+fit_perp[1,1]
AIC_BM<-2*1-2*logLike_BM
print(AIC_BM)

logLike_OU<-fit_par[1,2]+fit_perp[1,2]
AIC_OU<-2*2-2*logLike_OU
print(AIC_OU)
```

c) Run model with overlap proportion as continuous variable
```{r}
models = c("BM_null", "BM_linear", "OU_null", "OU_linear","OU_linear_beta")
e9 <- evodat$trait
t9<-evodat$p_distance_100
g9<-evodat$overlap_proportion
fit_par_patry <- model.test.sisters(e9,t9,g9,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t9,e9,xlim=c(0,10),ylim=c(0,1),main="no divide")
print(fit_par_patry) #View model parameters
```

d) Run model with separate rates for migration category AND including overlap proportion
```{r}
## parallels
evodat_parallel_patry<-subset(evodat,evodat$migration_category=="parallel")
e9 <- evodat_parallel_patry$trait
t9<-evodat_parallel_patry$p_distance_100
g9<-evodat_parallel_patry$overlap_proportion
fit_par_patry <- model.test.sisters(e9,t9,g9,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t9,e9,xlim=c(0,10),ylim=c(0,1),main="no divide")
print(fit_par_patry) #View model parameters

# ci for parallels
# intercept_beta_par<-as.numeric(fit_par_patry[5,4])
# slope_beta_par<-as.numeric(fit_par_patry[6,4])
# intercept_alpha_par<-as.numeric(fit_par_patry[11,4])
# slope_alpha_par<-as.numeric(fit_par_patry[12,4])
# parameters_par=c(intercept_beta_par, slope_beta_par, intercept_alpha_par, slope_alpha_par)
# set.seed(seed = 3)
# parR <- bootstrap.test(e9,t9,g9,model="OU_linear", parameters_par, meserr1=0, meserr2=0, breakpoint = "NULL", N = c(1000), starting=NULL)
# parR$summary

## perpendiculars
evodat_perp_patry<-subset(evodat,evodat$migration_category=="perpendicular")
e10 <- evodat_perp_patry$trait
t10 <-evodat_perp_patry$p_distance_100
g10 <-evodat_perp_patry$overlap_proportion
fit_perp_patry <- model.test.sisters(e10,t10,g10,GRAD2=NULL,meserr1=0,meserr2=0,models,starting=NULL,Beta_starting=NULL,Alpha_starting=NULL)
#plot(t10,e10,xlim=c(0,10),ylim=c(0,1),main="divide")
print(fit_perp_patry) #View model parameters

# ci for perpendiculars
# intercept_beta_perp<-as.numeric(fit_perp_patry[5,4])
# slope_beta_perp<-as.numeric(fit_perp_patry[6,4])
# intercept_alpha_perp<-as.numeric(fit_perp_patry[11,4])
# slope_alpha_perp<-as.numeric(fit_perp_patry[12,4])
# parameters_perp=c(intercept_beta_perp, slope_beta_perp, intercept_alpha_perp, slope_alpha_perp)
# set.seed(seed = 3)
# perpR <- bootstrap.test(e10,t10,g10,model="OU_linear", parameters_perp, meserr1=0, meserr2=0, breakpoint = "NULL", N = c(1000), starting=NULL)
# perpR$summary

## calculate AIC for perp/par with patry and using _linear models
logLike_BM<-fit_par_patry[1,2]+fit_perp_patry[1,2]
AIC_BM<-2*1-2*logLike_BM
print(AIC_BM)

logLike_OU<-fit_par_patry[1,4]+fit_perp_patry[1,4]
AIC_OU<-2*2-2*logLike_OU
print(AIC_OU)

## calculate AIC for perp/par with patry and using _linear_beta models
logLike_OU<-fit_par_patry[1,5]+fit_perp_patry[1,5]
AIC_OU<-2*2-2*logLike_OU
print(AIC_OU)
```

